---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
fig.path = "man/figures/README-",
out.width = "100%"
)
```

# IncidencePrevalence

<!-- badges: start -->
[![codecov.io](https://codecov.io/github/darwin-eu/IncidencePrevalence/coverage.svg?branch=main)](https://codecov.io/github/darwin-eu/IncidencePrevalence?branch=main)
[![R-CMD-check](https://github.com/darwin-eu/IncidencePrevalence/workflows/R-CMD-check/badge.svg)](https://github.com/darwin-eu/IncidencePrevalence/actions)
[![Lifecycle:Experimental](https://img.shields.io/badge/Lifecycle-Experimental-339999)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

## Package overview

IncidencePrevalence contains functions for estimating population-level incidence and prevalence using the OMOP common data model.  


## Package installation

You can install the latest version of IncidencePrevalence like so:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("darwin-eu/IncidencePrevalence")
```

## Example usage
### Create a reference to data in the OMOP CDM format 
The IncidencePrevalence package is designed to work with data in the OMOP CDM format, so our first step is to create a reference to the data using the CDMConnector package. Here we´ll generate an example reference with simulated data (to see how you would create a reference to your database please consult the CDMConnector package documentation).

```{r}
library(CDMConnector)
library(IncidencePrevalence)

# We first need to create a cdm_reference 
cdm<-mockIncidencePrevalenceRef(sampleSize=5000)
# this is what the person table for this example data looks like
head(cdm$person)
```

### Identify a denominator cohort
To identify a set of denominator cohorts we can use the `generateDenominatorCohortSet` function. Here we want to identify denominator populations for a study period between 2008 and 2018 and with 180 days of prior history (observation time in the database). We also wish to consider multiple age groups (from 0 to 64, and 65 to 100) and multiple sex criteria (one cohort only males, one only females, and one with both sexes included).  

```{r}
cdm$denominator <- generateDenominatorCohortSet(
  cdm = cdm,
  startDate = as.Date("2008-01-01"),
  endDate = as.Date("2018-01-01"),
  ageGroups = list(c(0,64),
                   c(65,100)),
  sex = c("Male", "Female", "Both"),
  daysPriorHistory = 180
)
# this is what the resulting cohort data looks like
head(cdm$denominator)
```

This will then give us six denominator cohorts
```{r}
settings(cdm$denominator)
```

For each of these cohorts, we can see the reasons people in the database did not enter a given cohort 
```{r}
attrition(cdm$denominator)
```

### Estimating incidence and prevalence
Now we have identified our denominator population, we can calculate incidence and prevalence as below. Note, in our example cdm reference we already have an outcome cohort defined as for this package outcome cohorts are required to have been defined externally.

For this example we´ll estimate incidence on a yearly basis, allowing individuals to have multiple events but with an outcome washout of 180 days. We also require that only complete database intervals are included, by which we mean that the database must have individuals observed throughout a year for that year to be included in the analysis. Note, we also specify a minimum cell count of 5, under which estimates will be obscured.

```{r}
inc <- estimateIncidence(
  cdm = cdm,
  denominatorTable = "denominator",
  outcomeTable = "outcome",
  interval = "years",
  repeatedEvents = TRUE,
  outcomeWashout = 180, 
  completeDatabaseIntervals = TRUE,
  minCellCount = 5
  )
dplyr::glimpse(inc)
```

We could also estimate point prevalence, as of the start of each calendar year like so:
```{r}
prev_point <- estimatePointPrevalence(
  cdm = cdm,
  denominatorTable = "denominator",
  outcomeTable = "outcome",
  interval = "years",
  timePoint = "start",
  minCellCount = 5
)
dplyr::glimpse(prev_point)
```

And annual period prevalence where we again require complete database intervals and, in addition, only include those people who are observed in the data for the full year:

```{r}
prev_period <- estimatePeriodPrevalence(
  cdm = cdm,
  denominatorTable = "denominator",
  outcomeTable = "outcome",
  interval = "years",
  completeDatabaseIntervals = TRUE, 
  fullContribution = TRUE,
  minCellCount = 5
)
dplyr::glimpse(prev_period)
```

### Combining and exporting results
After running different analyses We can use `gatherResults()` to bring together the results, adding outcome names and the database name to the output.
```{r}
study_results <- gatherResults(resultList=list(inc, prev_point, prev_period),
                    outcomeCohortId = 1,
                    outcomeCohortName = "example_outcome",
                    databaseName = "example_data")
dplyr::glimpse(study_results$incidence_estimates)
dplyr::glimpse(study_results$prevalence_estimates)
```

After gathering results, we can export them as CSVs in a zip folder using the `writeResultToDisk()` function. 
```{r, eval=FALSE}
writeResultToDisk(result=study_results, 
                  zipName="example_results",
                  outputFolder=here::here()) 
```
