---
title: "PaRe Report: IncidencePrevalence [0.3.1]"
author: "`r paste('PaRe', packageVersion('PaRe'))`"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
vignette: >
  %\VignetteIndexEntry{IncidencePrevalence [0.3.1]}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(PaRe)
library(dplyr)
```

```{r, echo=FALSE}
repo <- PaRe::Repository$new(here::here())
```

## Package overview
```{r packageDiagram, out.width="100%", warning=FALSE, echo=FALSE}
repo %>%
  PaRe::pkgDiagram() %>%
  DiagrammeRsvg::export_svg() %>%
  charToRaw() %>%
  magick::image_read()
```

## Summary of package functions
```{r getDefinedFunctionsPkg, message=FALSE, warning=FALSE, echo=FALSE}
defFuns <- PaRe::getDefinedFunctions(repo)

defFuns <- defFuns %>%
  dplyr::mutate(size = lineEnd - lineStart + 1)
```

```{r summaryDefFunStats, echo=FALSE}
DT::datatable(
  data.frame(
    min = c(
      min(defFuns$nArgs, na.rm = TRUE),
      min(defFuns$size, na.rm = TRUE),
      min(defFuns$cycloComp, na.rm = TRUE)
    ),
    median = c(
      median(defFuns$nArgs, na.rm = TRUE),
      median(defFuns$size, na.rm = TRUE),
      median(defFuns$cycloComp, na.rm = TRUE)
    ),
    max = c(
      max(defFuns$nArgs, na.rm = TRUE),
      max(defFuns$size, na.rm = TRUE),
      max(defFuns$cycloComp, na.rm = TRUE)
    ),
    row.names = c(
      "Number of arguments",
      "Lines of code",
      "Cyclomatic complexity"
    )
  ),
  options = list(dom = "t")
)
```

```{r echo=FALSE}
gg_nArgs <- defFuns %>%
  ggplot2::ggplot() +
  ggplot2::geom_histogram(ggplot2::aes(nArgs),
    binwidth = 1,
    colour = "black",
    fill = "grey"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::xlab("Number of arguments")

gg_size <- defFuns %>%
  ggplot2::ggplot() +
  ggplot2::geom_histogram(ggplot2::aes(size),
    binwidth = 1,
    colour = "black",
    fill = "grey"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::xlab("Lines of code")

gg_cycloComp <- defFuns %>%
  ggplot2::ggplot() +
  ggplot2::geom_histogram(ggplot2::aes(cycloComp),
    binwidth = 1,
    colour = "black",
    fill = "grey"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::xlab("Cyclomatic complexity")

cowplot::plot_grid(gg_nArgs, gg_size, gg_cycloComp, nrow = 3)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
p <- defFuns %>%
  ggplot2::ggplot(ggplot2::aes(group = name)) +
  ggplot2::geom_point(
    ggplot2::aes(size, cycloComp,
      colour = nArgs
    ),
    size = 3
  ) +
  ggplot2::scale_colour_gradient(
    name = "Number of\nargruments",
    low = "blue", high = "red"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::xlab("Lines of code") +
  ggplot2::ylab("Cyclomatic complexity") +
  ggplot2::theme(legend.position = "top")

plotly::ggplotly(p,
  tooltip = c(
    "group", "colour",
    "x", "y"
  )
)
```

## Function details
```{r definedFunctions, echo=FALSE}
colours <- c("", "#81d4fa", "#03a9f4")
argBreaks <- c(5, 7)
complexBreaks <- c(15, 30)
lineBreaks <- c(100, 200)

DT::datatable(
  defFuns %>%
    dplyr::mutate(file_start = paste0(.data$fileName, " (from line: ", lineStart, ")")) %>%
    dplyr::select("name", "nArgs", "size", "cycloComp", "file_start") %>%
    dplyr::rename(
      "Function" = "name",
      "Number of arguments" = "nArgs",
      "Lines of code" = "size",
      "Cyclomatic complexity" = "cycloComp",
      "Location" = "file_start"
    ),
  rownames = FALSE
) %>%
  DT::formatStyle("Number of arguments", backgroundColor = DT::styleInterval(argBreaks, colours)) %>%
  DT::formatStyle("Cyclomatic complexity", backgroundColor = DT::styleInterval(complexBreaks, colours)) %>%
  DT::formatStyle("Lines of code", backgroundColor = DT::styleInterval(lineBreaks, colours))
```

## Use of other programming languages
```{r countLines, echo=FALSE}
DT::datatable(
  PaRe::countPackageLines(repo),
  rownames = "# lines of code"
)
```

## Style adherence, syntax errors and possible semantic issues
### Summary
```{r lintScores, message=FALSE, warning=FALSE, echo=FALSE}
lintMsgs <- PaRe::lintRepo(repo)

DT::datatable(
  PaRe::lintScore(repo, lintMsgs) %>%
    dplyr::rename("Percentage of lines assessed" = "pct"),
  rownames = FALSE
)
```

### Warnings
```{r lintMessages warnings, message=FALSE, warning=FALSE, echo=FALSE}
DT::datatable(
  lintMsgs %>%
    dplyr::filter(type == "warning") %>%
    dplyr::group_by(message) %>%
    dplyr::tally(sort = TRUE),
  rownames = FALSE
)
```

### Style
```{r lintMessages style, message=FALSE, warning=FALSE, echo=FALSE}
DT::datatable(
  lintMsgs %>%
    dplyr::filter(type == "style") %>%
    dplyr::group_by(message) %>%
    dplyr::tally(sort = TRUE),
  rownames = FALSE
)
```

## Dependency review
### Check dependencies against whitelist
```{r checkDependencies, warning=FALSE, echo=FALSE}
DT::datatable(PaRe::checkDependencies(repo = repo, verbose = FALSE))
```

### Dependency characteristics
- **countVertices**: The amount of recursive dependencies `r repo$getName()` depends on.
- **countEdges**: The total amount of imports of all dependencies.
- **meanDegree**: The average amount of imports per dependency.
- **meanDistance**: The average dependency layers between `r repo$getName()` and all other recursive dependencies.
```{r dependencyGraphStats, message=FALSE, warning=FALSE, echo=FALSE}
graphData <- PaRe::getGraphData(repo = repo)

DT::datatable(
  data.frame(
    countVertices = length(igraph::V(graphData)),
    countEdges = length(igraph::E(graphData)),
    meanDegree = round(mean(igraph::degree(graphData)), 2),
    meanDistance = round(mean(igraph::distances(graphData)), 2)
  ),
  rownames = FALSE
)
```

### Function use per dependency
```{r summariseFunctionUse, message=FALSE, warning=FALSE, echo=FALSE}
funsUsed <- PaRe::getFunctionUse(repo = repo)

DT::datatable(
  funsUsed,
  rownames = FALSE
)
```

```{r plotFunctionUse, dpi=100, fig.height=25, out.width="100%", message=FALSE, warning=FALSE, echo=FALSE}
function_sub <- funsUsed %>%
  dplyr::filter(!pkg %in% c("base"))

fun_counts <- function_sub %>%
  dplyr::group_by(fun, pkg, name = "n") %>%
  dplyr::tally()

# Filter functions that are defined in the package
nonPkgFuns <- fun_counts[!fun_counts$fun %in% defFuns$fun, ]

ggplot2::ggplot(
  data = nonPkgFuns,
  mapping = ggplot2::aes(x = .data$fun, y = .data$n, fill = .data$pkg)
) +
  ggplot2::geom_col() +
  ggplot2::facet_wrap(
    dplyr::vars(.data$pkg),
    scales = "free_x",
    ncol = 2
  ) +
  ggplot2::theme_bw() +
  ggplot2::theme(
    legend.position = "none",
    axis.text.x = (ggplot2::element_text(angle = 45, hjust = 1, vjust = 1))
  )
```

## Further reading
### Style Guide
The styling is based on the following style guide:

- [Tidyverse styleguide](https://style.tidyverse.org/syntax.html)
- [HADES styleguide](https://ohdsi.github.io/Hades/codeStyle.html)

### Cyclomatic Complexity
>[Cyclomatic Complexity](https://en.wikipedia.org/wiki/Cyclomatic_complexity) is a software metric used to indicate the complexity of a program. It is a quantitative measure of the number of linearly independent paths through a program's source code. It was developed by Thomas J. McCabe, Sr. in 1976.

And is calculated as follows: $M = E - N + 2P$

The complexity score is interpreted as follows:

> 
> - 1-10 Simple procedure, little risk
> - 11-20 More complex, moderate risk
> - 21-50 Complex, high risk
> - \>50 Untestable code, very high risk
> 

[link, 2023-02-10](https://en.wikipedia.org/wiki/Cyclomatic_complexity#Interpretation)
